### 2024-07-11 analysis/preprocessing.qmd
minor updates morning

## Summarise

@@

### Pull data to analyse regression studies

@@

## Export as csv

@@

>>>>

## Summarise all

@@

## Pull data to analyse regression studies

@@

### Create descriptive table for the manuscript

```{r}
TR <- NULL
TR$study_n <- length(unique(R_studies$citekey))
TR$model_n <- nrow(R_studies)
t<-table(R_studies$model_class_id)
t2 <- paste0(names(t),': ', as.numeric(t))
TR$model_types_n <- str_c(t2,collapse = "\n")
TR$feature_Desc <- paste0('Min=',min(R_studies$feature_n,na.rm = TRUE),', Md=',median(R_studies$feature_n,na.rm = TRUE),', Max=', max(R_studies$feature_n,na.rm = TRUE))
TR$stimulus_Desc <- paste0('Min=',min(R_studies$stimulus_n,na.rm = TRUE),', Md=',median(R_studies$stimulus_n,na.rm = TRUE),', Max=', max(R_studies$stimulus_n,na.rm = TRUE))
print(TR)

### Export as csv

@@

## Pull data to analyse classification studies

```{r}
C_studies <- dplyr::filter(metaMER_results,model_category=='classification' & str_detect(measure,'accuracy|r2|mcc')) #  
dim(C_studies)

# eliminate measures that we don't need now
C_studies <- dplyr::filter(C_studies,values>0)
C_studies$statistic[is.na(C_studies$statistic)]<-'Not defined'
C_studies <- dplyr::filter(C_studies,!str_detect(statistic,'pvalue|upper|lower'))
C_studies <- dplyr::filter(C_studies,str_detect(measure,'accuracy|mcc|r2'))
dim(C_studies)

table(C_studies$measure,C_studies$statistic)

dim(C_studies)
# Conversion into mcc (==r), do we need

```{r}
#C_studies$dimension[str_detect(R_studies$dimension,'activation|energy arousal|tension arousal')]<-'arousal'
#R_studies$dimension[str_detect(R_studies$dimension,'pleasantness')]<-'valence'
#R_studies <- dplyr::filter(R_studies,!str_detect(dimension,'av')) # relates to distances, can be omitted

table(C_studies$stimulus_n)

# Deal with four studies involving multiple datasets: 
C_studies$stimulus_n[C_studies$stimulus_n==" 429; 350 popular songs + 79 songs from the Beatles (Mirex 2009 collection)  "] <- 429 # resolved from the paper
C_studies$stimulus_n[C_studies$stimulus_n==" 5192; 12 per user in user validation (not included here due to little information),   AcousticBrainz validation: 60000  "] <- 5192 # resolved
C_studies$stimulus_n[C_studies$stimulus_n==" ISMIR2012: 2886, NJU_V1: 777, Hindi: 1037  "] <- 2886+777+1037 # decided to take this from validation
C_studies<-C_studies[!str_detect(C_studies$stimulus_n,'not specified'),]
C_studies<-C_studies[!str_detect(C_studies$stimulus_n,'Specified elsewhere'),]
C_studies$stimulus_n <- as.numeric(C_studies$stimulus_n)

### Create descriptive table for the manuscript

```{r}
TR <- NULL
TR$study_n <- length(unique(R_studies$citekey))
TR$model_n <- nrow(R_studies)
t<-table(R_studies$model_class_id)
t2 <- paste0(names(t),': ', as.numeric(t))
TR$model_types_n <- str_c(t2,collapse = "\n")
TR$feature_Desc <- paste0('Min=',min(R_studies$feature_n,na.rm = TRUE),', Md=',median(R_studies$feature_n,na.rm = TRUE),', Max=', max(R_studies$feature_n,na.rm = TRUE))
TR$stimulus_Desc <- paste0('Min=',min(R_studies$stimulus_n,na.rm = TRUE),', Md=',median(R_studies$stimulus_n,na.rm = TRUE),', Max=', max(R_studies$stimulus_n,na.rm = TRUE))
print(TR)

### Export as csv

### 2024-07-11 manuscript/manuscript.qmd
minor updates morning

- Great advances in music information retrieval have been made in recent years.

@@

- Our aim is to establish the level of predictive accuracy for both continuous models of emotional expression (valence and arousal) and classification of emotion categories based on available and recent studies. 
- We seek to identify the types of issues (modelling techniques, features, and musical qualities used) that significantly influence the prediction rates.
- To achieve these aims, we carry out a meta-analysis focused on journal articles published in the last 10 years.
- We outline broad hypotheses such as arousal being predicted to a higher degree than valence, which is more challenging and more context dependent than arousal. For classification, simple utilitarian emotions (e.g., fear, anger) will be easier to predict than complex social emotions (e.g., sadness, nostalgia).

@@

We preregistered our meta-analysis plan (OSF URL here, also available at [https://tuomaseerola.github.io/metaMER/preregistration/preregistration.html](https://tuomaseerola.github.io/metaMER/preregistration/preregistration.html)).

@@

@@

First we could describe the overall pattern of data (modelling techniques, cross-validation, popular datasets, etc).

@@

## Prediction success for valence and arousal

Out of XX studies reporting this task, the majority (xx%) ...

| Concept | N    | Md $R^2$|
|:--------|:-----|:--------|
| Valence | ??   | ??      |
| Arousal | ??   | ??      |

TABLE: xxxx

@@

To be done.

>>>>

- Great advances in music information retrieval have been made in recent years. The available features, modelling techniques and datasets have given scholars opportunities to refine the accuracy and reliability of predicting annotated emotions from audio.

@@

- Our aim is to establish the level of predictive accuracy for both models of emotional expression that can account for track-specific coordinates in affective circumple space (valence and arousal) and classification of emotion categories based on available and recent studies. 
- We seek to identify the types of issues (modelling techniques, features, and musical qualities used) that significantly influence the prediction rates.
- To achieve these aims, we carry out a meta-analysis focused on journal articles published in the last 10 years.
- We outline broad hypotheses such as arousal being predicted to a higher degree than valence, which is more challenging and more context dependent than arousal. For classification, simple utilitarian emotions (e.g., fear, anger) will be easier to predict than complex social emotions (e.g., sadness, nostalgia).

@@

We preregistered our meta-analysis plan on 21 June 2024 at OSF, [https://osf.io/c5wgd](https://osf.io/c5wgd), and the plan is also available at [https://tuomaseerola.github.io/metaMER/preregistration/preregistration.html](https://tuomaseerola.github.io/metaMER/preregistration/preregistration.html)).

@@

[optional Figure: flowchart of the study inclusions/eliminations]

- Devote some sentences to quality control issues (what were the most common reasons for eliminating studies).

@@

First we could describe the overall pattern of data (regression vs classification, modelling techniques, feature numbers, stimulus numbers, datasets, and other details).

TABLE 1: Summary of data (part of `analysis/preprocessing.qmd`)

Talk about the stimuli and

@@

## Prediction success for valence and arousal (or affect dimensions)

See `analysis/analysis.qmd`

- Decision: Report all models or the best model?
  * Possible solution: Report all models, trimmed models, and but mainly rely on Max

- Report issues of heterogeneity right here (can be visualised with funnel plots, see below)

- Optional Figure X: Funnel plot that combines valence and arousal (separate panels) with FULL/TRIMMED data shown with markers"

Table 2. Meta-analytic diagnostic for all regression studies predicting valence from audio.

| Concept     | Models, obs|$r$ [95%-CI]              | $t$ | $p$  |$\tau^2$|$I^2$|
|:------------|:----------|:--------|:--------------|:----|:-----|:-----|:------|
| Valence All | 49,42648  | 0.6137 [0.5540; 0.6672] | 15.9|.0001 | 0.092| 98.1% |
| Valence Trim| 27,18402  | 0.6048 [0.5920; 0.6172] | 72.4|.0001 | 0.0007| 33.1% [0.0%; 58.2%]|
| Valence M   | 14,10641  | 0.6038 [0.4671; 0.7124] | 7.83|.0001 | 0.100| 97.8% |
| Valence Md  | 14,10641  | 0.6018 [0.4659; 0.7099] | 7.87|.0001 | 0.098| 97.7% |
| Valence Max | 14,10641  | 0.6520 [0.5026; 0.7636] | 7.44|.0001 | 0.142| 97.9% |
| *N Features*|           |                         |     |      |      |       |
| <18 F       |           | 0.xxxx [0.xxxx; 0.xxxx] | xxxx|.xxxx | 0.xxx| XXXX% |
| 18-260 F    |           | 0.xxxx [0.xxxx; 0.xxxx] | xxxx|.xxxx | 0.xxx| XXXX% |
| 260+ F      |           | 0.xxxx [0.xxxx; 0.xxxx] | xxxx|.xxxx | 0.xxx| XXXX% |
| *Techniques*|           |                         |     |      |      |       |
|   KS        |           | 0.3606 [0.3201; 0.3997] | xxxx|.xxxx | 0.xxx| XXXX% |
|   LM        |           | 0.xxxx [0.xxxx; 0.xxxx] | xxxx|.xxxx | 0.xxx| XXXX% |
|   FD        |           | 0.xxxx [0.xxxx; 0.xxxx] | xxxx|.xxxx | 0.xxx| XXXX% |
|   NN        |           | 0.xxxx [0.xxxx; 0.xxxx] | xxxx|.xxxx | 0.xxx| XXXX% |
|   RF        |           | 0.xxxx [0.xxxx; 0.xxxx] | xxxx|.xxxx | 0.xxx| XXXX% |

Figure 1. Forest plot of valence prediction (Max?)
`analysis/figure1.qmd`

Summarise the results here briefly

Moving on the arousal, ...

Table 3. Meta-analytic diagnostic for all regression studies predicting arousal from audio.

| Concept     | Models, obs|$r$ [95%-CI]             | $t$ | $p$  |$\tau^2$|$I^2$|
|:------------|:-----------|:---------|:-------------|:----|:-----|:-----|:------|
| Arousal     | 14,9035    |          |              |     |      |      |       |
| Arousal Trim| 27,18402   |          |              |     |      |      |       |
| Arousal M   | 14,10641   |          |              |     |      |      |       |
| Arousal Md  | 14,10641   |          |              |     |      |      |       |
| Arousal Max | 14,10641   |          |              |     |      |      |       |
| *N Features*|            |          |              |     |      |      |       |   
| <18 F       |            |          |              |     |      |      |       |
| 18-260 F    |            |          |              |     |      |      |       |
| 260+ F      |            |          |              |     |      |      |       |
| *Techniques*|            |          |              |     |      |      |       |  
|   KS        |            |          |              |     |      |      |       |
|   LM        |            |          |              |     |      |      |       |
|   FD        |            |          |              |     |      |      |       |
|   NN        |            |          |              |     |      |      |       |
|   RF        |            |          |              |     |      |      |       |

Figure 2. Forest plot of arousal prediction (using Max?)

Summarise here the pattern of results

## Classification studies

Summary of details contained in Table 1, but summarise at least the categories predicted before moving onto the main findings.

Table 4. Meta-analytic diagnostic for all classification studies predicting emotion categories from audio.

Heterogeneity issues

Figure 3. Forest plot of arousal prediction (Max?) (Unless we do some custom plotting)

- Figure Optional: Funnel plot (I haven't seen this yet)

@@

## Concise summary of what we did and found

## Main outcomes 
- Arousal is easier to predict (r = 0.7627) than valence (r = 0.6236), as we predicted. The glass ceiling seems to be at ...
- Classification ...
- Model accuracy is surprisingly little affected by the number of features (?) or modelling technique (?). 
- Some of the complex state-of-the-art techniques (e.g., NNs) do not deliver impressive improvements over older techniques (e.g., SVR, RF)
- Variation in study/model/data quality is large and can be seen in heterogenuity and the amount of studies eliminated

## Calls for action/points to improve in such studies
- *Documentation* the details in full (features, stimuli, model details, cross-validation)
- *Quality* of the underlying data (emotion ratings, classes, or even stimulus properties?
- *Generalisibility* of the models (some studies such as X and Y address this by applying the models across several datasets)
- Diversity in the evaluative aspects of studies: *overfitting*, numerous ways of cross-validating, not sharing data or analysis scripts, not reporting in the same way
- What proportion of stimuli are Western music, and what genres tend to dominate?
